---
Title: Assessment c7081; Who eats what we grow
Author: "mayowa-Adebayo.R"
Date: "2022-11-22"
Output: EDA using logical regression
output: word_document
---

## Header ####

## Who: "22338000"

## What: "Assessment c7081; Who eats what we grow"

## Last Edited: "2022-11-22"

## What: "Exploratory Data Analysis using logical regresssion"

## Github link: [https://github.com/mayowa-data/c7081]

---

## Content ####
       
# 01.Background
# 02.Methodology
## i. Set working directory
## ii. Load necessary library
## iii. Read in dataset
## iv. Cleanup/tidy dataset
##  v. Analyse data using logical regression
#  03. Result 
#  04. Conclusion
#  05. References

---

## 01 Background

Who eats what we grow
           
Link:[https://www.kaggle.com/datasets/dorbicycle/world-foodfeed-production] 

             
The population of the world is estimated to grow from 8billion today
to 9.7billion in the year 2050.(fischer et al., 2002).
The panacea of providing food to these ever growing population has become major 
concern for FAO, Government, Entrepreneur, and philanthropist. (FAO., 2015)

There are various ways to combat these challenges and some of the ways include;
Improving on our agricultural practices as well as adjusting our diets. Things
gets even more complicated with climate change as it affects food production - 
Agriculture.

This dataset gives us perspective on the food production worldwide. However,
our focus in this analysis is based on food/ feed production in the UK from
2003 - 2013. The question is in line with the article " Who eats the food we grow

Our Question: Which was produced more in the united kingdom between 2003 and 2013
Food or Feed.

The dataset was from Kaggle.

## 02 Methodology

 The dataset was extracted from kaggel dataset on "Who eat what we grow", 
 compiled and published by the Food and Agriculture Organization
 of the united nation.
 The original dataset consist of Food product data from 1961 to 2013 
 from 245 countries and territories. The focus of this analysis however, is to 
 determine which is produced more in the united kingdom between 
 the period of 2003 to 2013; Feed for animals or food for humans.
 The logic in this question poses a classification problem and we shall be using
 logistic regression  to solve the problem; 
 because dependent variable is either Feed or Food.

```{r}
# set working directory

setwd("C:/Users/MAYOWA/OneDrive/Desktop/Scripts")

library(openxlsx)

foodfeed_production <- read.xlsx("Feedfoodproduction.xlsx")

sum(is.na(foodfeed_production))
foodfeed_production <- na.omit(foodfeed_production) #remove missing data
class <- foodfeed_production[ , 7] #class label
foodfeed_production2 <- foodfeed_production[ , 53:63]
foodfeed_production3 <- cbind(class, foodfeed_production2) #combine class label+
                                                          #descriptorinto same df
df<- foodfeed_production3 #df is now the new dataframe
df$class <- as.factor(df$class) #convert to factor

```


```{r}
#cor() produce matrix that shows the pairwise correlation between the varriable

cor(df[ , -1]) #1st variable is the class containing food or feed.
```


```{r}
# We use glm() for logistic regression.

glm.fit<- glm(class ~ Y2003+Y2004+Y2005+Y2006+Y2007+Y2008+Y2009+Y2010+
                Y2011+Y2012+Y2013, data = df, family = binomial)

summary(glm.fit)
```

```{r}
#coef() determine the coefficients of glm.fit

coef(glm.fit)
```


```{r}
#predict() for probability of producing Food or Feed

glm.prob <- predict(glm.fit, type = "response")
glm.prob[1:10]
```

```{r}

#convert predicted probability to label "Food or Feed"

contrasts(df$class)
```



```{r}
#Create a vector of class prediction on probability of 
#production will be greater or less than 0.8

glm.pred <- rep("Feed", 17938) # make vector
glm.pred[glm.prob > .8] <- "Food" # add Food values

```

```{r}
#determine which of the observation is correctly of incorrectly classified
table(glm.pred, df$class) # confusion matrix

```


```{r}
mean(glm.pred == class) # to determine the proportion of correct prediction
```


```{r}
#set training model (create a vector corresponding to observation dividing
#the data into 70% and 30%
#training model 70%.
#test model 30%

set.seed(100)
train_index <- sample(2,nrow(df), replace = TRUE, prob =c(0.7,0.30)) #training data
train.data <- df[train_index== 1,]
df.Y2013 <- df[train_index==2,]
dim(df.Y2013)
```


```{r} 
#test model
class.Y2013 <- class[ train_index==2]
```



```{r}
#fit logistic regression to subset of test model

glm.fit <- glm(class ~ Y2003 +Y2004+Y2005+Y2006+Y2007+Y2008+Y2009+Y2010+
                Y2011+Y2012+Y2013, data = train.data,
              family = binomial)
predict.probs <- predict(glm.fit, df.Y2013, type = "response")

```


```{r}
#compute the prediction for test model and compare it with training model
glm.pred <- rep("Feed", 5272)
glm.pred[predict.probs > .8] <- "Food"

table(glm.pred, class.Y2013)
```


```{r}
#mean of the predicted test model
mean(glm.pred == class.Y2013)

```


```{r}
#mean of not ! the predicted test model
mean(glm.pred != class.Y2013)
```


```{r}
#plot the glm.fit graph

par(mfrow = c(2,2))
plot(glm.fit)
```


## 03 Result
The result of our above analysis shows that 83% of the time, 
our prediction was correct that more food was produced in the United Kingdom
between the year 2003 and 2013. The test error scores shows 17% which inferred
that the error in our prediction is low. 


## 04 Conclusion
We can conclude from our analysis that more food was produced in 
the United Kingdom in the period in view(2003- 2013). which means that 
more food were cultivated for human consumption than for animal feed. Humans are 
the main consumer of What we grow.


## 05 Reference

Fischer, G., Shah, M.M., & van Velthuizen, H.T. (2002). 
Climate Change and Agricultural Vulnerability.
IIASA, Laxenburg, Austria.

Food and Agriculture Organization 2015
UN DESA report, “World Population Prospects"
New York

Gareth James, Daniela Witten , Trevor Hastie, Robert Tibshirani
An Introduction to Statistical Learning 
with Applications in R
Second Edition
First printing : August 4, 2021


